% Created 2020-11-15 Sun 16:59
% Intended LaTeX compiler: pdflatex
\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{grffile}
\usepackage{longtable}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{textcomp}
\usepackage{amssymb}
\usepackage{capt-of}
\usepackage{hyperref}
\usepackage{amsthm}
\usepackage{minted}
\author{Tan Yee Jian}
\date{\today}
\title{Assignment 7 Report}
\hypersetup{
 pdfauthor={Tan Yee Jian},
 pdftitle={Assignment 7 Report},
 pdfkeywords={},
 pdfsubject={},
 pdfcreator={Emacs 27.0.91 (Org mode 9.4)}, 
 pdflang={English}}
\begin{document}

\maketitle
\tableofcontents

This is the last assignment for this module, and last report from me. Good job
everyone for coming this far!

Everyone had some understanding for OOP, but not everyone can produce the most
succinct and least error-prone code. Only a couple have gotten full marks. To
get full marks, you must use existing code as much as possible, as copy-pasting
is always a big source for error.

As for inheritance, please remember that child classes inherit all functions by
default. There is no need to override and use \texttt{super().act()} in the body
\textbf{only} - it is redundant code.
\section{Berserker}
\label{sec:org2fa93b9}
We discussed this question in tutorial, and I apologize in advance if I did not
stress that you will need to make sure that strength will return to normal when
HP is back up. Other than that, everyone got this part.
\section{Archmage}
\label{sec:org4f3908c}
Most people forgot to check if the opponent is alive before using \texttt{got\_hurt()}.
Other than that, a good number edited the \texttt{act()} function instead of the
\texttt{cast()} function and got penalized. I feel this is a reasonable one since
\texttt{KABOOM} is intuitively a spell, and using \texttt{cast()} is just natural.

When you are doing normal attacks, do not copy the code, but instead, use
\texttt{super().cast()}.
\section{Necromancer}
\label{sec:orga4854e6}
This part sees more success than the previous parts, despite having higher
marks. Some of you used \texttt{count\_alive(team)!=0} and I would suggest using
\texttt{> 0} instead. This can sieve out weird bugs such as, for some reason,
\texttt{count\_alive()} was wrongly implemented.
\end{document}
