@inproceedings{10.1145/2213836.2213862,
  author = {Sears, Russell and Ramakrishnan, Raghu},
  title = {BLSM: A General Purpose Log Structured Merge Tree},
  year = {2012},
  isbn = {9781450312479},
  publisher = {Association for Computing Machinery},
  address = {New York, NY, USA},
  url = {https://doi-org.libproxy1.nus.edu.sg/10.1145/2213836.2213862},
  doi = {10.1145/2213836.2213862},
  abstract = {Data management workloads are increasingly write-intensive and subject to strict latency
    SLAs. This presents a dilemma: Update in place systems have unmatched latency but
    poor write throughput. In contrast, existing log structured techniques improve write
    throughput but sacrifice read performance and exhibit unacceptable latency spikes.We
    begin by presenting a new performance metric: read fanout, and argue that, with read
    and write amplification, it better characterizes real-world indexes than approaches
    such as asymptotic analysis and price/performance.We then present bLSM, a Log Structured
    Merge (LSM) tree with the advantages of B-Trees and log structured approaches: (1)
    Unlike existing log structured trees, bLSM has near-optimal read and scan performance,
    and (2) its new "spring and gear" merge scheduler bounds write latency without impacting
    throughput or allowing merges to block writes for extended periods of time. It does
    this by ensuring merges at each level of the tree make steady progress without resorting
    to techniques that degrade read performance.We use Bloom filters to improve index
    performance, and find a number of subtleties arise. First, we ensure reads can stop
    after finding one version of a record. Otherwise, frequently written items would incur
    multiple B-Tree lookups. Second, many applications check for existing values at insert.
    Avoiding the seek performed by the check is crucial.},
  booktitle = {Proceedings of the 2012 ACM SIGMOD International Conference on Management of Data},
  pages = {217–228},
  numpages = {12},
  keywords = {merge scheduling, read amplification, read fanout, write amplification, log structured merge tree},
  location = {Scottsdale, Arizona, USA},
  series = {SIGMOD '12}
}
